% ============================================================================
%  Industrial DMZ Design - Poster / Cheat Sheet
% ============================================================================

\documentclass[9pt,a4paper]{extarticle}
\usepackage{otsec-poster}
\usepackage{float}

\begin{document}

\makepostertitle
    {Industrial DMZ Design}
    {Secure Buffer Zone Architecture for OT Networks}
    {Poster 302}
    {Matthias Niedermaier}

\begin{multicols}{2}

\section{\textcolor{accent}{\faIcon{server}}\hspace{0.4em}Overview}

An Industrial DMZ (IDMZ) is a network segment acting as a \textbf{secure buffer between enterprise IT and OT networks}. Provides controlled connectivity while preventing direct communication between business and control systems. Sits at \textbf{Purdue Level 3.5}.

\posterinfo{
\faIcon{shield-alt}\hspace{0.2em}The IDMZ is one of the most critical security controls in OT. A properly designed DMZ prevents lateral movement from compromised IT systems to critical control networks.
}

\section{\textcolor{accent}{\faIcon{columns}}\hspace{0.4em}IT DMZ vs Industrial DMZ}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2cm}p{2cm}p{2.5cm}}
\rowcolor{primary}
\textcolor{white}{\bfseries Aspect} & \textcolor{white}{\bfseries IT DMZ} & \textcolor{white}{\bfseries IDMZ} \\
\midrule
\faIcon{bullseye}\hspace{0.2em}Purpose & Expose to internet & Separate IT/OT \\
\faIcon{exchange-alt}\hspace{0.2em}Traffic & Inbound & Bidirectional \\
\faIcon{cubes}\hspace{0.2em}Services & Web, email, DNS & Historians, jump \\
\faIcon{handshake}\hspace{0.2em}Trust & Internet untrusted & IT semi-trusted \\
\faIcon{clock}\hspace{0.2em}Availability & High & Critical \\
\end{tabular}
\end{center}

\section{\textcolor{accent}{\faIcon{project-diagram}}\hspace{0.4em}Architecture Patterns}

\subsection{\textcolor{accent}{\faIcon{shield-alt}}\hspace{0.3em}Dual-Firewall DMZ (Recommended)}

\begin{center}
\begin{tikzpicture}[
    zone/.style={rectangle, draw=#1!60, thick, fill=#1!10, rounded corners=3pt, minimum width=5.5cm, minimum height=0.5cm, align=center, font=\scriptsize},
    zone/.default={otaccent},
    fw/.style={rectangle, draw=otdanger!80, thick, fill=otdanger!20, minimum width=5.5cm, minimum height=0.35cm, align=center, font=\scriptsize\bfseries, text=otdanger},
]
    \node[zone=red] (ent) at (0,0) {\faIcon{building}\hspace{0.2em}Enterprise Network (L4--5)};
    \node[fw, below=3pt of ent] (ofw) {\faIcon{shield-alt}\hspace{0.2em}Outer Firewall (Vendor A)};
    \node[zone=orange, below=3pt of ofw] (dmz) {\faIcon{server}\hspace{0.2em}Industrial DMZ (L3.5)};
    \node[fw, below=3pt of dmz] (ifw) {\faIcon{shield-alt}\hspace{0.2em}Inner Firewall (Vendor B)};
    \node[zone=green!60!black, below=3pt of ifw] (ot) {\faIcon{industry}\hspace{0.2em}OT Network (L0--3)};

    \draw[->, thick, otaccent] (ent) -- (ofw);
    \draw[->, thick, otaccent] (ofw) -- (dmz);
    \draw[->, thick, otaccent] (dmz) -- (ifw);
    \draw[->, thick, otaccent] (ifw) -- (ot);
\end{tikzpicture}
\end{center}

\postersuccess{
\faIcon{check-circle}\hspace{0.2em}\textbf{Different vendors} for inner/outer firewalls prevents single vulnerability from compromising both. OT initiates outbound connections to DMZ.
}

\subsection{\textcolor{accent}{\faIcon{long-arrow-alt-right}}\hspace{0.3em}Data Diode Integration}

For highest security, add hardware-enforced unidirectional gateways:
\begin{itemize}
    \item \faIcon{arrow-right}\hspace{0.2em}\textbf{Outbound diode:} OT to DMZ (historian replication)
    \item \faIcon{ban}\hspace{0.2em}\textbf{No inbound path:} Physically prevents IT attacks
    \item \faIcon{industry}\hspace{0.2em}\textbf{Use cases:} Nuclear, critical infrastructure
\end{itemize}

\section{\textcolor{accent}{\faIcon{list-ul}}\hspace{0.4em}DMZ Services}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.5cm}p{4cm}}
\rowcolor{primary}
\textcolor{white}{\bfseries Service} & \textcolor{white}{\bfseries Purpose} \\
\midrule
\faIcon{database}\hspace{0.2em}Historian Mirror & Read-only process data for business \\
\faIcon{laptop}\hspace{0.2em}Jump Server & Secure remote access to OT \\
\faIcon{download}\hspace{0.2em}Patch Server & Tested update staging area \\
\faIcon{shield-alt}\hspace{0.2em}AV/Update Server & Signature and software updates \\
\faIcon{file-alt}\hspace{0.2em}File Transfer & Secure IT-OT file exchange \\
\faIcon{lock}\hspace{0.2em}RA Gateway & VPN termination for remote users \\
\faIcon{clipboard-list}\hspace{0.2em}Log Collector & OT security log aggregation \\
\end{tabular}
\end{center}

\posterwarning{
\faIcon{exclamation-triangle}\hspace{0.2em}\textbf{Service placement rules:} No service should have simultaneous connections to IT and OT. Data must ``break'' in the DMZ. OT initiates connections outbound. IT connects only to DMZ, never to OT.
}

\section{\textcolor{accent}{\faIcon{exchange-alt}}\hspace{0.4em}Traffic Flow Design}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{2cm}p{2.3cm}}
\rowcolor{primary}
\textcolor{white}{\bfseries Source} & \textcolor{white}{\bfseries Dest} & \textcolor{white}{\bfseries Purpose} \\
\midrule
\faIcon{industry}\hspace{0.2em}OT $\rightarrow$ DMZ & Hist mirror & Data replication \\
\faIcon{industry}\hspace{0.2em}OT $\rightarrow$ DMZ & Log collect & Security events \\
\faIcon{server}\hspace{0.2em}DMZ $\rightarrow$ OT & Patch srv & Tested updates \\
\faIcon{building}\hspace{0.2em}IT $\rightarrow$ DMZ & Hist mirror & Data queries \\
\faIcon{building}\hspace{0.2em}IT $\rightarrow$ DMZ & Jump srv & Remote admin \\
\end{tabular}
\end{center}

\posterdanger{
\textbf{Never allow:}
\begin{itemize}
    \item \textcolor{danger}{\faIcon{times}}\hspace{0.2em}IT directly to OT (any protocol)
    \item \textcolor{danger}{\faIcon{times}}\hspace{0.2em}OT directly to IT
    \item \textcolor{danger}{\faIcon{times}}\hspace{0.2em}Internet directly to DMZ or OT
    \item \textcolor{danger}{\faIcon{times}}\hspace{0.2em}Broad ``any-any'' rules through the DMZ
\end{itemize}
}

\section{\textcolor{accent}{\faIcon{filter}}\hspace{0.4em}Firewall Configuration}

\begin{enumerate}
    \item \faIcon{ban}\hspace{0.2em}\textbf{Default deny} -- block all not explicitly permitted
    \item \faIcon{key}\hspace{0.2em}\textbf{Least privilege} -- only required ports/protocols
    \item \faIcon{crosshairs}\hspace{0.2em}\textbf{Specific sources/dest} -- no ``any'' in rules
    \item \faIcon{search}\hspace{0.2em}\textbf{Application awareness} -- OT-aware DPI
    \item \faIcon{clipboard-list}\hspace{0.2em}\textbf{Logging} -- log all traffic, especially denied
\end{enumerate}

\section{\textcolor{accent}{\faIcon{eye}}\hspace{0.4em}Monitoring and Management}

\begin{itemize}
    \item \faIcon{chart-line}\hspace{0.2em}\textbf{Traffic analysis:} Monitor for anomalous patterns
    \item \faIcon{shield-alt}\hspace{0.2em}\textbf{IDS/IPS:} Deploy at DMZ boundaries
    \item \faIcon{database}\hspace{0.2em}\textbf{Log correlation:} Aggregate from all DMZ systems
    \item \faIcon{bell}\hspace{0.2em}\textbf{Alerting:} Immediate notification of violations
    \item \faIcon{tasks}\hspace{0.2em}\textbf{Change management:} Strict change control
    \item \faIcon{flask}\hspace{0.2em}\textbf{Testing:} Validate changes in non-production first
    \item \faIcon{users}\hspace{0.2em}\textbf{Approval:} Both IT and OT stakeholders
\end{itemize}

\postertip{
\faIcon{lightbulb}\hspace{0.2em}The dual-firewall IDMZ with different vendor firewalls is the gold standard. OT systems should \textbf{push data outbound} to the DMZ -- never allow inbound connections from IT into OT. Quarterly rule reviews prevent ``rule creep.''
}

\end{multicols}

\end{document}
