% ============================================================================
%  OT Security Standards Overview - Poster / Cheat Sheet
% ============================================================================

\documentclass[9pt,a4paper]{extarticle}
\usepackage{otsec-poster}
\usepackage{float}

\begin{document}

\makepostertitle
    {OT Security Standards Overview}
    {Global Industrial Cybersecurity Regulations}
    {Poster 100}
    {Matthias Niedermaier}

\begin{multicols}{2}

\section{\textcolor{accent}{\faIcon{globe}}\hspace{0.4em}Overview}

The convergence of IT and OT, combined with rising cyber threats to critical infrastructure, has led to numerous cybersecurity standards worldwide. Understanding these is essential for multi-region or multi-sector organizations.

\section{\textcolor{accent}{\faIcon{sitemap}}\hspace{0.4em}Standards Landscape}

\begin{center}
\begin{tikzpicture}[
    std/.style={rectangle, draw=#1!60, thick, fill=#1!12, rounded corners=2pt,
        minimum height=0.4cm, align=center, font=\scriptsize},
    std/.default={otaccent},
    cat/.style={font=\scriptsize\bfseries, text=#1},
    conn/.style={-, thick, #1!30},
    conn/.default={otaccent},
]
    % Center - IEC 62443
    \node[std=otprimary, minimum width=2.2cm, minimum height=0.5cm, font=\scriptsize\bfseries] (iec) at (0,0) {IEC 62443};

    % International
    \node[std=otinfo, minimum width=1.5cm] (iso) at (-2.2,0.8) {ISO 27001};
    \node[std=otinfo, minimum width=1.5cm] (isa) at (2.2,0.8) {ISA-95};

    % Regional
    \node[std=otsuccess, minimum width=1.5cm] (nist) at (-2.2,-0.1) {NIST 800-82};
    \node[std=otdanger, minimum width=1.5cm] (nerc) at (-2.2,-0.9) {NERC CIP};
    \node[std=otwarning, minimum width=1.2cm] (nis2) at (2.2,-0.1) {NIS2};
    \node[std=otwarning, minimum width=1.2cm] (cra) at (2.2,-0.9) {CRA};

    % Connections
    \draw[conn=otinfo] (iso) -- (iec);
    \draw[conn=otinfo] (isa) -- (iec);
    \draw[conn=otsuccess] (nist) -- (iec);
    \draw[conn=otdanger] (nerc) -- (iec);
    \draw[conn=otwarning] (nis2) -- (iec);
    \draw[conn=otwarning] (cra) -- (iec);

    % Region labels
    \node[cat=otinfo] at (0,1.2) {International};
    \node[cat=otsuccess] at (-2.8,-0.5) {USA};
    \node[cat=otwarning] at (2.8,-0.5) {EU};
\end{tikzpicture}
\end{center}

\posterinfo{
\textbf{Types of standards:} International (ISO, IEC, ISA), Regional regulations (EU, North America), National guidelines, and Sector-specific (energy, water, manufacturing).
}

\section{\textcolor{accent}{\faIcon{map}}\hspace{0.4em}Global Standards Map}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2cm}p{2.8cm}p{1.8cm}}
\rowcolor{primary}
\textcolor{white}{\bfseries Region} & \textcolor{white}{\bfseries Key Standards} & \textcolor{white}{\bfseries Mandatory} \\
\midrule
International & IEC 62443, ISO 27001 & No \\
North America & NIST 800-82, NERC CIP & Partial \\
Europe & NIS2, CRA, BSI & Yes \\
Asia-Pacific & GB/T 30976, METI & Partial \\
Australia/NZ & SOCI Act, AESCSF & Yes \\
Middle East & NESA (UAE), NCA (SA) & Yes \\
\end{tabular}
\end{center}

\section{\textcolor{accent}{\faIcon{globe-americas}}\hspace{0.4em}International Standards}

\subsection{\textcolor{accent}{\faIcon{gavel}}\hspace{0.3em}IEC 62443}
Comprehensive IACS security framework. Four series: General, Policies/Procedures, System, Component. Covers asset owners, integrators, and vendors. \textbf{Most widely referenced standard globally.}

\subsection{\faIcon{certificate}\hspace{0.3em}ISO/IEC 27001}
Information security management systems (ISMS). Certifiable. General framework often combined with IEC 62443 for OT-specific requirements.

\subsection{\faIcon{sitemap}\hspace{0.3em}ISA-95 / IEC 62264}
Enterprise-control system integration. Defines functional hierarchy, basis for Purdue Model.

\section{\textcolor{accent}{\faIcon{flag-usa}}\hspace{0.4em}North America}

\begin{itemize}
    \item \textcolor{info}{\faIcon{book}}\hspace{0.2em}\textbf{NIST SP 800-82:} Primary US guidance for OT security (Rev. 3, 2023). Voluntary.
    \item \textcolor{danger}{\faIcon{exclamation-triangle}}\hspace{0.2em}\textbf{NERC CIP:} \textbf{Mandatory} for bulk electric system. Penalties up to \$1M/violation/day.
    \item \textcolor{accent}{\faIcon{shield-alt}}\hspace{0.2em}\textbf{NIST CSF:} Risk-based framework, all sectors.
    \item \textcolor{accent}{\faIcon{road}}\hspace{0.2em}\textbf{TSA Directives:} Pipeline and rail cybersecurity requirements.
    \item \textcolor{accent}{\faIcon{flag}}\hspace{0.2em}\textbf{Canada:} CSA Z246.1 (petroleum/gas), NERC CIP (power).
\end{itemize}

\section{\textcolor{accent}{\faIcon{globe}}\hspace{0.4em}Europe}

\posterdanger{
\textbf{NIS2 Directive:} Mandatory for essential and important entities across EU. Significantly expands scope -- energy, transport, health, water, digital infrastructure, manufacturing. Member states must transpose into national law.
}

\begin{itemize}
    \item \textcolor{danger}{\faIcon{shield-alt}}\hspace{0.2em}\textbf{Cyber Resilience Act (CRA):} Mandatory cybersecurity for products with digital elements -- hardware and software including OT components.
    \item \textcolor{accent}{\faIcon{building}}\hspace{0.2em}\textbf{Germany:} BSI IT-Grundschutz (OT module: IND), IT-Sicherheitsgesetz 2.0, KRITIS Verordnung.
    \item \textcolor{accent}{\faIcon{flag}}\hspace{0.2em}\textbf{UK:} NIS Regulations, NCSC Cyber Assessment Framework.
    \item \textcolor{accent}{\faIcon{flag}}\hspace{0.2em}\textbf{France:} LPM (Military Programming Law), ANSSI Guidelines.
\end{itemize}

\section{\textcolor{accent}{\faIcon{globe-asia}}\hspace{0.4em}Asia-Pacific \& Others}

\begin{itemize}
    \item \textcolor{accent}{\faIcon{flag}}\hspace{0.2em}\textbf{China:} GB/T 22239, GB/T 30976 (ICS), Cybersecurity Law (mandatory for CI).
    \item \textcolor{accent}{\faIcon{flag}}\hspace{0.2em}\textbf{Japan:} METI Cyber-Physical Security Framework, NISC.
    \item \textcolor{accent}{\faIcon{flag}}\hspace{0.2em}\textbf{Australia:} SOCI Act, AESCSF (energy), Essential Eight.
\end{itemize}

\section{\textcolor{accent}{\faIcon{industry}}\hspace{0.4em}Sector-Specific Standards}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{1.5cm}p{2.2cm}p{3cm}}
\rowcolor{primary}
\textcolor{white}{\bfseries Sector} & \textcolor{white}{\bfseries Standard} & \textcolor{white}{\bfseries Scope} \\
\midrule
\faIcon{bolt}\hspace{0.1em}Power & NERC CIP & Bulk electric system \\
\faIcon{gas-pump}\hspace{0.1em}Oil \& Gas & API 1164 & Pipeline SCADA \\
\faIcon{tint}\hspace{0.1em}Water & AWWA Guide & Water utilities \\
\faIcon{atom}\hspace{0.1em}Nuclear & NRC 10 CFR & US nuclear cybersecurity \\
\faIcon{ship}\hspace{0.1em}Maritime & IMO MSC.428 & Cyber risk management \\
\faIcon{train}\hspace{0.1em}Rail & EN 50129 & Safety-related systems \\
\faIcon{car}\hspace{0.1em}Auto & ISO/SAE 21434 & Vehicle cybersecurity \\
\end{tabular}
\end{center}

\section{\textcolor{accent}{\faIcon{columns}}\hspace{0.4em}Standards Comparison}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{1.8cm}p{0.7cm}p{0.7cm}p{0.7cm}p{2cm}}
\rowcolor{primary}
\textcolor{white}{\bfseries Standard} & \textcolor{white}{\bfseries Mand.} & \textcolor{white}{\bfseries Cert.} & \textcolor{white}{\bfseries OT} & \textcolor{white}{\bfseries Region} \\
\midrule
IEC 62443 & No & Yes & Yes & International \\
ISO 27001 & No & Yes & No & International \\
NIST 800-82 & No & No & Yes & USA \\
NERC CIP & Yes & No & Yes & N. America \\
NIS2 & Yes & No & Part & EU \\
GB/T 30976 & Yes & No & Yes & China \\
\end{tabular}
\end{center}

\posterwarning{
\textbf{Compliance is not security.} Standards provide a minimum baseline. Organizations should use risk-based approaches to go beyond checklists -- especially in OT where safety and availability are paramount.
}

\postertip{
When operating across multiple regions, use \textbf{IEC 62443 as a baseline} framework and map to regional requirements. This provides a common foundation while ensuring local compliance.
}

\end{multicols}

\end{document}
