% ============================================================================
%  Privileged Access Management - Poster / Cheat Sheet
% ============================================================================

\documentclass[9pt,a4paper]{extarticle}
\usepackage{otsec-poster}
\usepackage{float}

\begin{document}

\makepostertitle
    {Privileged Access Management}
    {Controlling Administrative Access in OT Environments}
    {Poster 880}
    {Matthias Niedermaier}

\begin{multicols}{2}

\section{\textcolor{accent}{\faIcon{info-circle}}\hspace{0.4em}Overview}

PAM controls and monitors access by users with elevated permissions. In OT, privileged accounts include admin credentials, \textbf{vendor remote access, engineering workstation logins}, and service accounts that can modify control system configurations.

\posterinfo{
\textbf{OT impact:} A compromised privileged account can modify PLC logic, disable safety systems, or disrupt physical processes. OT privileged access affects safety and operations, not just data.
}

\section{\textcolor{accent}{\faIcon{user-lock}}\hspace{0.4em}Privileged Account Types}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{user-lock}\hspace{0.2em}\bfseries Type} & \textcolor{white}{\faIcon{exclamation-triangle}\hspace{0.2em}\bfseries Description and Risk} \\
\midrule
System admin & Full access to servers, workstations \\
OT engineer & Can modify PLC/DCS logic, HMI configs \\
Vendor/Integrator & Remote maintenance, broad permissions \\
Service accounts & Automated processes, elevated privs \\
Break-glass & High-privilege emergency accounts \\
Shared accounts & Generic logins, multiple operators \\
\end{tabular}
\end{center}

\posterwarning{
\textbf{OT challenges:} Legacy systems lack modern auth. Shared accounts on HMIs. Vendor access poorly controlled. 24/7 operations need always-on access. Safety systems require rapid access. Air-gapped systems complicate central PAM.
}

\section{\textcolor{accent}{\faIcon{project-diagram}}\hspace{0.4em}PAM Architecture}

\subsection{\textcolor{accent}{\faIcon{cogs}}\hspace{0.3em}Core Components}

\begin{itemize}
    \item \textcolor{accent}{\faIcon{lock}}\hspace{0.2em}\textbf{Credential vault} -- Store and rotate privileged credentials
    \item \textcolor{accent}{\faIcon{desktop}}\hspace{0.2em}\textbf{Session gateway} -- Proxy and record sessions
    \item \textcolor{accent}{\faIcon{clipboard-check}}\hspace{0.2em}\textbf{Access request workflow} -- Manage approval process
    \item \textcolor{accent}{\faIcon{clock}}\hspace{0.2em}\textbf{Just-in-time access} -- Grant temporary elevation
    \item \textcolor{accent}{\faIcon{video}}\hspace{0.2em}\textbf{Session recording} -- Capture keystrokes, screens, commands
\end{itemize}

\subsection{\textcolor{accent}{\faIcon{server}}\hspace{0.3em}Jump Server Architecture}

All administrative access to OT should traverse hardened jump servers in the DMZ:

\begin{itemize}
    \item \textcolor{accent}{\faIcon{cog}}\hspace{0.2em}Minimal software, full session logging
    \item \textcolor{accent}{\faIcon{fingerprint}}\hspace{0.2em}Multi-factor authentication required
    \item \textcolor{accent}{\faIcon{ban}}\hspace{0.2em}Direct connections to OT systems blocked
    \item \textcolor{accent}{\faIcon{layer-group}}\hspace{0.2em}Separate jump servers per security zone
\end{itemize}

\section{\textcolor{accent}{\faIcon{truck}}\hspace{0.4em}Vendor Access Management}

\posterdanger{
\textbf{Vendor remote access is a top OT attack vector.} Attackers compromise vendor credentials or infrastructure, then use trusted connections to reach customer sites. The Kaseya attack demonstrated this at scale.
}

\subsection{\textcolor{accent}{\faIcon{key}}\hspace{0.3em}Vendor Access Controls}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{key}\hspace{0.2em}\bfseries Control} & \textcolor{white}{\faIcon{cogs}\hspace{0.2em}\bfseries Implementation} \\
\midrule
No persistent & Disable connections when not in use \\
Explicit approval & Customer approves each session \\
Time-limited & Auto-terminate after defined period \\
Least privilege & Access only specific systems needed \\
MFA & Required for all vendor connections \\
Session recording & Record all vendor activity for audit \\
Real-time monitor & Alert on unusual vendor behavior \\
\end{tabular}
\end{center}

\subsection{\textcolor{accent}{\faIcon{tasks}}\hspace{0.3em}Vendor Access Workflow}

\begin{enumerate}
    \item \textcolor{accent}{\faIcon{paper-plane}}\hspace{0.2em}Vendor requests access with justification
    \item \textcolor{accent}{\faIcon{check-circle}}\hspace{0.2em}Customer approves and defines scope
    \item \textcolor{accent}{\faIcon{key}}\hspace{0.2em}Temporary credentials issued via PAM
    \item \textcolor{accent}{\faIcon{eye}}\hspace{0.2em}Monitored and recorded session
    \item \textcolor{accent}{\faIcon{times-circle}}\hspace{0.2em}Access automatically revoked after session
\end{enumerate}

\section{\textcolor{accent}{\faIcon{lock}}\hspace{0.4em}Credential Management}

\subsection{\textcolor{accent}{\faIcon{lock}}\hspace{0.3em}Password Vaulting}

\begin{itemize}
    \item \textcolor{accent}{\faIcon{lock}}\hspace{0.2em}Store all privileged credentials in encrypted vault
    \item \textcolor{accent}{\faIcon{sign-out-alt}}\hspace{0.2em}Users check out credentials through PAM system
    \item \textcolor{accent}{\faIcon{sync-alt}}\hspace{0.2em}Automatic rotation after each use or on schedule
    \item \textcolor{accent}{\faIcon{clipboard-list}}\hspace{0.2em}Full audit trail for all credential access
\end{itemize}

\subsection{\textcolor{accent}{\faIcon{sync-alt}}\hspace{0.3em}Rotation Challenges in OT}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{exclamation-triangle}\hspace{0.2em}\bfseries Challenge} & \textcolor{white}{\faIcon{check-circle}\hspace{0.2em}\bfseries Mitigation} \\
\midrule
Hardcoded creds & Plan remediation in maintenance \\
Service account deps & Map dependencies before rotation \\
Air-gapped systems & Manual rotation with verification \\
Legacy (no APIs) & Agent-based or scripted rotation \\
24/7 operations & Rotate during planned maintenance \\
\end{tabular}
\end{center}

\section{\textcolor{accent}{\faIcon{eye}}\hspace{0.4em}Session Management}

\begin{itemize}
    \item \textcolor{accent}{\faIcon{video}}\hspace{0.2em}\textbf{Video recording} -- Capture screen activity for GUI
    \item \textcolor{accent}{\faIcon{keyboard}}\hspace{0.2em}\textbf{Keystroke logging} -- Record all commands entered
    \item \textcolor{accent}{\faIcon{filter}}\hspace{0.2em}\textbf{Command filtering} -- Block or alert on dangerous commands
    \item \textcolor{accent}{\faIcon{bell}}\hspace{0.2em}\textbf{Real-time alerts} -- Notify on high-risk activities
\end{itemize}

\section{\textcolor{accent}{\faIcon{exclamation-circle}}\hspace{0.4em}Emergency Access}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{exclamation-circle}\hspace{0.2em}\bfseries Element} & \textcolor{white}{\faIcon{info-circle}\hspace{0.2em}\bfseries Description} \\
\midrule
Emergency accts & Pre-created, high-privilege, sealed \\
Notification & Auto-alert when emergency used \\
Time limits & Auto-disable after period \\
Post-review & Mandatory review of all use \\
\end{tabular}
\end{center}

\postertip{
Focus initial PAM on highest-risk access: vendor remote access, domain admin/root accounts, engineering workstations, then service accounts. Force all access through monitored jump servers. Record all sessions. Implement approval workflows for vendor access with time limits and auto-revocation.
}

\end{multicols}

\end{document}
