% ============================================================================
%  Industroyer/CrashOverride - Poster / Cheat Sheet
% ============================================================================

\documentclass[9pt,a4paper]{extarticle}
\usepackage{otsec-poster}
\usepackage{float}

\begin{document}

\makepostertitle
    {Industroyer / CrashOverride}
    {Analysis of Grid-Targeting Malware}
    {Poster 413}
    {Matthias Niedermaier}

\begin{multicols}{2}

\section{\textcolor{accent}{\faIcon{info-circle}}\hspace{0.4em}Overview}

Industroyer (CrashOverride) is sophisticated malware designed to attack \textcolor{danger}{\faIcon{industry}}\hspace{0.2em}\textbf{electric power grids}. Used in the December 2016 Ukraine attack, it caused a blackout in Kyiv affecting $\sim$200MW of load. Only the second known malware (after Stuxnet) specifically designed to disrupt physical industrial processes.

\section{\textcolor{accent}{\faIcon{list-ul}}\hspace{0.4em}Incident Facts}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{list-ul}\hspace{0.2em}\bfseries Attribute} & \textcolor{white}{\faIcon{info-circle}\hspace{0.2em}\bfseries Details} \\
\midrule
Date & December 17, 2016 \\
Target & \textcolor{danger}{\faIcon{industry}}\hspace{0.2em}Ukrenergo (Kyiv transmission) \\
Location & Pivnichna substation \\
Impact & $\sim$200MW, 1-hour outage \\
Attribution & Sandworm (Russian GRU) \\
Discovery & ESET and Dragos (June 2017) \\
\end{tabular}
\end{center}

\section{\textcolor{accent}{\faIcon{cubes}}\hspace{0.4em}Modular Architecture}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{cubes}\hspace{0.2em}\bfseries Component} & \textcolor{white}{\faIcon{cogs}\hspace{0.2em}\bfseries Function} \\
\midrule
\textcolor{warning}{\faIcon{bug}}\hspace{0.2em}Main backdoor & C2 communication, orchestration \\
\textcolor{warning}{\faIcon{bug}}\hspace{0.2em}Add. backdoor & Persistence mechanism \\
Launcher & Schedule-based payload execution \\
Protocol mods & ICS protocol attack components \\
\textcolor{warning}{\faIcon{bug}}\hspace{0.2em}Data wiper & Evidence destruction, damage \\
\end{tabular}
\end{center}

\section{\textcolor{accent}{\faIcon{stream}}\hspace{0.4em}ICS Protocol Modules}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{stream}\hspace{0.2em}\bfseries Protocol} & \textcolor{white}{\faIcon{crosshairs}\hspace{0.2em}\bfseries Attack Capability} \\
\midrule
IEC 60870-5-101 & Serial comms with RTUs \\
IEC 60870-5-104 & TCP/IP RTU comms (primary) \\
IEC 61850 & Substation auto., GOOSE \\
OPC DA & SCADA historian/HMI access \\
\end{tabular}
\end{center}

\posterdanger{
\textbf{Four industrial protocol modules} show extensive power grid knowledge. The modular design allows rapid adaptation to different targets worldwide. Fully automated attack requiring no operator interaction.
}

\section{\textcolor{accent}{\faIcon{crosshairs}}\hspace{0.4em}Attack Sequence}

\begin{center}
\begin{tikzpicture}[
    step/.style={rectangle, draw=#1!40, thick, fill=#1!5, minimum height=0.38cm, rounded corners=2pt, font=\scriptsize, text width=4.5cm, align=left},
    step/.default={otdanger},
    num/.style={circle, fill=#1, text=white, font=\scriptsize\bfseries, minimum size=0.33cm, inner sep=0pt},
    num/.default={otdanger},
]
    \node[num=otwarning] (n1) at (0,0) {1};
    \node[step=otwarning, right=3pt of n1] (s1) {Spearphishing / supply chain};
    \node[num=otwarning, below=2pt of n1] (n2) {2};
    \node[step=otwarning, right=3pt of n2] (s2) {Lateral move through enterprise};
    \node[num=otdanger, below=2pt of n2] (n3) {3};
    \node[step=otdanger, right=3pt of n3] (s3) {Pivot to control system network};
    \node[num=otdanger, below=2pt of n3] (n4) {4};
    \node[step=otdanger, right=3pt of n4] (s4) {Recon and deploy malware};
    \node[num=otdanger, below=2pt of n4] (n5) {5};
    \node[step=otdanger, right=3pt of n5] (s5) {\faIcon{bolt}\hspace{0.1em}Open breakers (23:50 local)};
    \node[num=otprimary, below=2pt of n5] (n6) {6};
    \node[step=otprimary, right=3pt of n6] (s6) {\faIcon{eraser}\hspace{0.1em}Data wiper destroys evidence};

    \foreach \i/\j in {1/2,2/3,3/4,4/5,5/6} {
        \draw[thick, otdanger!40] (n\i.south) -- (n\j.north);
    }
\end{tikzpicture}
\end{center}

\subsection{\textcolor{accent}{\faIcon{bolt}}\hspace{0.3em}IEC 104 Attack Details}

\begin{itemize}
    \item \textcolor{warning}{\faIcon{search}}\hspace{0.2em}Scanned network for IEC 104 devices
    \item \textcolor{warning}{\faIcon{list-alt}}\hspace{0.2em}Enumerated Information Object Addresses (IOAs)
    \item \textcolor{danger}{\faIcon{keyboard}}\hspace{0.2em}Sent unauthorized commands to open breakers
    \item \textcolor{danger}{\faIcon{exchange-alt}}\hspace{0.2em}Issued single-point and double-point commands
    \item \textcolor{danger}{\faIcon{sync}}\hspace{0.2em}Toggled breakers repeatedly to prevent manual restore
\end{itemize}

\section{\textcolor{accent}{\faIcon{columns}}\hspace{0.4em}Comparison: 2015 vs 2016}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{1.8cm}p{2.2cm}p{2.5cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{columns}\hspace{0.2em}\bfseries Aspect} & \textcolor{white}{\bfseries 2015} & \textcolor{white}{\bfseries 2016} \\
\midrule
Method & Remote desktop & \textcolor{warning}{\faIcon{bug}}\hspace{0.2em}Automated malware \\
Protocols & Proprietary & IEC 104, 61850 \\
Scale & 225,000 affected & Smaller impact \\
Reusability & Low (manual) & High (modular) \\
Sophistic. & Medium & High \\
\end{tabular}
\end{center}

\posterinfo{
Industroyer represents an \textbf{evolution from manual to automated attacks}. The modular design allows rapid adaptation to different targets. This is a reusable weapon, not a one-time tool.
}

\section{\textcolor{accent}{\faIcon{cogs}}\hspace{0.4em}Technical Details}

\subsection{\textcolor{accent}{\faIcon{bug}}\hspace{0.3em}Main Backdoor}

\begin{itemize}
    \item \textcolor{warning}{\faIcon{bug}}\hspace{0.2em}Written in C++, compiled for Windows
    \item \textcolor{warning}{\faIcon{globe}}\hspace{0.2em}HTTPS to C2 servers, optional Tor anonymization
    \item \textcolor{warning}{\faIcon{upload}}\hspace{0.2em}File upload/download, shell commands
    \item \textcolor{warning}{\faIcon{file-code}}\hspace{0.2em}Configurable via XML configuration files
\end{itemize}

\subsection{\textcolor{accent}{\faIcon{eraser}}\hspace{0.3em}Data Wiper}

\begin{itemize}
    \item \textcolor{danger}{\faIcon{eraser}}\hspace{0.2em}Overwrites files with random data
    \item \textcolor{danger}{\faIcon{cog}}\hspace{0.2em}Targets Windows registry and system files
    \item \textcolor{danger}{\faIcon{power-off}}\hspace{0.2em}Makes systems unbootable
    \item \textcolor{danger}{\faIcon{search}}\hspace{0.2em}Hampers forensic investigation
\end{itemize}

\subsection{\textcolor{accent}{\faIcon{search}}\hspace{0.3em}Detection Indicators}

\begin{itemize}
    \item \textcolor{info}{\faIcon{chart-bar}}\hspace{0.2em}Unusual IEC 104 traffic patterns or command sequences
    \item \textcolor{info}{\faIcon{plug}}\hspace{0.2em}Unauthorized connections to RTU/IED addresses
    \item \textcolor{info}{\faIcon{calendar-alt}}\hspace{0.2em}Suspicious scheduled tasks on SCADA systems
    \item \textcolor{info}{\faIcon{user-secret}}\hspace{0.2em}Tor network connections from OT systems
    \item \textcolor{info}{\faIcon{file-alt}}\hspace{0.2em}File modifications in system directories
\end{itemize}

\section{\textcolor{accent}{\faIcon{shield-alt}}\hspace{0.4em}Defense Measures}

\postersuccess{
\begin{itemize}
    \item \textcolor{success}{\faIcon{project-diagram}}\hspace{0.2em}Network segmentation between IT and OT
    \item \textcolor{success}{\faIcon{shield-alt}}\hspace{0.2em}ICS-aware intrusion detection systems
    \item \textcolor{success}{\faIcon{check-circle}}\hspace{0.2em}Application whitelisting on SCADA servers
    \item \textcolor{success}{\faIcon{eye}}\hspace{0.2em}Monitor IEC 104/61850 protocol traffic
    \item \textcolor{success}{\faIcon{hdd}}\hspace{0.2em}Offline backups of critical configurations
    \item \textcolor{success}{\faIcon{clipboard-list}}\hspace{0.2em}Incident response plans for \textcolor{danger}{\faIcon{industry}}\hspace{0.2em}grid emergencies
\end{itemize}
}

\postertip{
\textbf{Protocol-specific protections:} Deploy IEC 62351 security extensions. Implement command validation against expected operations. Use rate limiting to detect rapid command sequences. Restrict which systems can send control commands. Automation increases risk -- malware attacks faster than humans respond. Manual restoration procedures remain essential for recovery.
}

\end{multicols}

\end{document}
