% ============================================================================
%  OT Security Operations Center - Poster / Cheat Sheet
% ============================================================================

\documentclass[9pt,a4paper]{extarticle}
\usepackage{otsec-poster}
\usepackage{float}

\begin{document}

\makepostertitle
    {OT Security Operations Center}
    {Designing and Operating a SOC for Industrial Environments}
    {Poster 560}
    {Matthias Niedermaier}

\begin{multicols}{2}

\section{\textcolor{accent}{\faIcon{info-circle}}\hspace{0.4em}Overview}

A Security Operations Center (SOC) provides centralized monitoring, detection, and response. Extending SOC capabilities to OT requires \textbf{different approaches, skills, and tools} than traditional IT security operations, with emphasis on \textcolor{danger}{\faIcon{hard-hat}}\hspace{0.2em}operational safety and availability.

\section{\textcolor{accent}{\faIcon{building}}\hspace{0.4em}SOC Models for OT}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{1.5cm}p{1.7cm}p{1.7cm}p{1.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{th-list}\hspace{0.2em}\bfseries Aspect} & \textcolor{white}{\bfseries Dedicated} & \textcolor{white}{\bfseries Converged} & \textcolor{white}{\bfseries Hybrid} \\
\midrule
Expertise & Deep OT & Limited & Tiered \\
Cost & High & Lower & Medium \\
Coverage & 24/7 hard & 24/7 easy & 24/7 ok \\
Best for & Large OT & Small OT & Most orgs \\
\end{tabular}
\end{center}

\postersuccess{
\textbf{Recommendation:} Most organizations benefit from a \textbf{hybrid model}---IT SOC provides 24/7 monitoring with OT-trained analysts who escalate to OT specialists for investigation and response.
}

\section{\textcolor{accent}{\faIcon{users}}\hspace{0.4em}Staffing and Skills}

\subsection{\textcolor{accent}{\faIcon{graduation-cap}}\hspace{0.3em}Required Competencies}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{3cm}p{3.5cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{desktop}\hspace{0.2em}\bfseries IT Security} & \textcolor{white}{\faIcon{industry}\hspace{0.2em}\bfseries OT-Specific} \\
\midrule
Network analysis & Industrial protocols \\
Malware analysis & Control system architecture \\
SIEM operations & Process understanding \\
Incident response & Safety awareness \\
\end{tabular}
\end{center}

\subsection{\textcolor{accent}{\faIcon{layer-group}}\hspace{0.3em}Staffing Tiers}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{1.2cm}p{1.5cm}p{3.5cm}}
\rowcolor{primary}
\textcolor{white}{\bfseries Tier} & \textcolor{white}{\faIcon{user-tag}\hspace{0.2em}\bfseries Role} & \textcolor{white}{\bfseries OT Requirements} \\
\midrule
Tier 1 & Triage & Basic OT awareness, protocol ID \\
Tier 2 & Investigate & OT network, ICS malware \\
Tier 3 & Response & Deep OT, vendor coordination \\
SME & Expert & Control system engineering \\
\end{tabular}
\end{center}

\posterwarning{
\textbf{Talent scarcity:} OT security talent is scarce. Cross-train IT analysts in OT fundamentals and partner with OT engineering teams for deep expertise during incidents.
}

\section{\textcolor{accent}{\faIcon{server}}\hspace{0.4em}Technology Stack}

\subsection{\textcolor{accent}{\faIcon{cogs}}\hspace{0.3em}Core Components}

\begin{itemize}
    \item \faIcon{eye}\hspace{0.2em}\textbf{OT Network Monitoring} -- DPI for industrial protocols
    \item \faIcon{chart-bar}\hspace{0.2em}\textbf{SIEM/SOAR} -- Correlation, automation, dashboards
    \item \faIcon{list-alt}\hspace{0.2em}\textbf{Asset Inventory} -- Automated OT device tracking
    \item \faIcon{bug}\hspace{0.2em}\textbf{Vulnerability Management} -- OT-aware assessment
    \item \faIcon{broadcast-tower}\hspace{0.2em}\textbf{Threat Intelligence} -- ICS-specific feeds and IOCs
    \item \faIcon{key}\hspace{0.2em}\textbf{Secure Remote Access} -- Monitored vendor access
    \item \faIcon{desktop}\hspace{0.2em}\textbf{EDR/EPP} -- Endpoint protection where possible
\end{itemize}

\subsection{\textcolor{accent}{\faIcon{puzzle-piece}}\hspace{0.3em}Integration Challenges}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{exclamation-triangle}\hspace{0.2em}\bfseries Challenge} & \textcolor{white}{\bfseries Approach} \\
\midrule
Protocol support & Ensure SIEM parses Modbus, DNP3 \\
Data volume & Filter/aggregate at source \\
Network isolation & Data diodes, secure forwarders \\
Real-time & Tune for OT use cases \\
\end{tabular}
\end{center}

\section{\textcolor{accent}{\faIcon{eye}}\hspace{0.4em}Data Sources and Visibility}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{1.2cm}p{5.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{sitemap}\hspace{0.2em}\bfseries Zone} & \textcolor{white}{\bfseries Log Sources} \\
\midrule
DMZ & Jump servers, firewalls, historians \\
Zone 3 & SCADA servers, HMIs, eng. workstations \\
Zone 1-2 & Network TAPs, managed switches, PLCs* \\
\end{tabular}
\end{center}

\vspace{0.2em}
{\scriptsize *Limited logging capability on most controllers}

\posterdanger{
Many OT devices cannot run agents or forward logs. \textcolor{info}{\faIcon{network-wired}}\hspace{0.2em}\textbf{Network traffic analysis} is often the only visibility into Level 0--1 activity. Deploy monitoring at strategic points to maximize coverage.
}

\section{\textcolor{accent}{\faIcon{bell}}\hspace{0.4em}Detection Use Cases}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{search}\hspace{0.2em}\bfseries Use Case} & \textcolor{white}{\bfseries Detection Logic} \\
\midrule
Unauth. eng. access & PLC commands from non-eng stations \\
New OT device & ARP/DHCP for unknown MACs \\
Protocol anomaly & Invalid function codes, malformed pkts \\
Lateral movement & IT protocols (SMB, RDP) in OT \\
Config change & Write to PLCs outside change windows \\
Remote access abuse & VPN/jump server at unusual times \\
\end{tabular}
\end{center}

\subsection{\textcolor{accent}{\faIcon{sort-amount-down}}\hspace{0.3em}Alert Prioritization}

\begin{itemize}
    \item \textcolor{danger}{\faIcon{exclamation-circle}}\hspace{0.2em}\textbf{Critical} -- Safety system alerts, immediate escalation
    \item \faIcon{cog}\hspace{0.2em}\textbf{High} -- Control system changes, verify with operations
    \item \faIcon{chart-line}\hspace{0.2em}\textbf{Medium} -- Network anomalies, investigate within shift
    \item \faIcon{clipboard-list}\hspace{0.2em}\textbf{Low} -- Policy violations, queue for review
\end{itemize}

\section{\textcolor{accent}{\faIcon{clipboard-list}}\hspace{0.4em}Operations and Processes}

\subsection{\textcolor{accent}{\faIcon{book}}\hspace{0.3em}Runbooks and Playbooks}

\begin{itemize}
    \item \faIcon{level-up-alt}\hspace{0.2em}\textbf{Escalation paths} -- Include OT engineering contacts
    \item \faIcon{hard-hat}\hspace{0.2em}\textbf{Response constraints} -- Document safe vs. risky actions
    \item \faIcon{users}\hspace{0.2em}\textbf{Coordination} -- When to involve plant operations
    \item \faIcon{bullhorn}\hspace{0.2em}\textbf{Communication} -- Who to notify, in what order
\end{itemize}

\subsection{\textcolor{accent}{\faIcon{chart-bar}}\hspace{0.3em}Key Performance Indicators}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{tachometer-alt}\hspace{0.2em}\bfseries Metric} & \textcolor{white}{\bfseries Description} \\
\midrule
MTTD & Time from event to SOC awareness \\
MTTR & Time from detection to response \\
Alert volume & Distribution across Purdue levels \\
FP rate & Percentage of non-true incidents \\
Asset coverage & OT assets with monitoring visibility \\
\end{tabular}
\end{center}

\postertip{
Start with a \textbf{hybrid IT/OT SOC model} and build OT expertise over time. Deploy network traffic analysis where endpoint logging is unavailable. Build OT-specific detection rules---IT patterns don't translate directly. Coordinate alert thresholds with operations and integrate with change management. OT SOC analysts must understand operational context: a traffic spike during batch processing is normal, at 3 AM it's suspicious.
}

\end{multicols}

\end{document}
