% ============================================================================
%  OT Tabletop Exercises - Poster / Cheat Sheet
% ============================================================================

\documentclass[9pt,a4paper]{extarticle}
\usepackage{otsec-poster}
\usepackage{float}

\begin{document}

\makepostertitle
    {OT Tabletop Exercises}
    {Testing Incident Response Through Scenario-Based Discussion}
    {Poster 650}
    {Matthias Niedermaier}

\begin{multicols}{2}

\section{\textcolor{accent}{\faIcon{info-circle}}\hspace{0.4em}Overview}

Tabletop exercises are discussion-based sessions where participants walk through simulated incident scenarios to \textbf{test response plans, identify gaps, and improve coordination}. For OT, these are essential because real-world testing on production systems is often impossible without risking safety.

\posterinfo{
\textbf{Why tabletops for OT?} Unlike IT where systems can be taken offline for testing, OT runs continuously. Tabletop exercises provide a safe way to validate incident response without disrupting operations or creating safety hazards.
}

\section{\textcolor{accent}{\faIcon{list-ul}}\hspace{0.4em}Exercise Types}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{list-ul}\hspace{0.2em}\bfseries Type} & \textcolor{white}{\faIcon{align-left}\hspace{0.2em}\bfseries Description} \\
\midrule
Tabletop & Facilitated discussion, low cost/risk \\
Workshop & Collaborative planning session \\
Drill & Tests specific procedure/capability \\
Functional & Multi-team operational exercise \\
Full-scale & End-to-end, high cost/complexity \\
\end{tabular}
\end{center}

\section{\textcolor{accent}{\faIcon{clipboard-list}}\hspace{0.4em}Planning and Preparation}

\subsection{\textcolor{accent}{\faIcon{bullseye}}\hspace{0.3em}Define Objectives}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{bullseye}\hspace{0.2em}\bfseries Objective} & \textcolor{white}{\faIcon{pencil-alt}\hspace{0.2em}\bfseries Example} \\
\midrule
Test procedures & Validate OT IR plan steps \\
Identify gaps & Find missing playbooks \\
Train personnel & Familiarize with escalation \\
Test comms & Verify notification chains \\
Assess decisions & Evaluate isolation criteria \\
\end{tabular}
\end{center}

\subsection{\textcolor{accent}{\faIcon{chess}}\hspace{0.3em}Scenario Design}

Effective OT scenarios should be:

\begin{itemize}
    \item \textcolor{success}{\faIcon{check-circle}}\hspace{0.2em}\textbf{Realistic} -- Based on actual threats to your industry
    \item \textcolor{info}{\faIcon{crosshairs}}\hspace{0.2em}\textbf{Relevant} -- Target systems participants manage
    \item \textcolor{warning}{\faIcon{brain}}\hspace{0.2em}\textbf{Challenging} -- Force difficult decisions
    \item \textcolor{accent}{\faIcon{hourglass-half}}\hspace{0.2em}\textbf{Scoped} -- Achievable within the exercise timeframe
\end{itemize}

\postersuccess{
\textbf{Scenario tip:} Base scenarios on real incidents from your sector. Adapt published case studies (Stuxnet, TRITON, Colonial Pipeline, Oldsmar) to your specific environment and systems.
}

\subsection{\textcolor{accent}{\faIcon{tasks}}\hspace{0.3em}Preparation Checklist}

\begin{enumerate}
    \item \faIcon{file-alt}\hspace{0.2em}Scenario narrative and injects
    \item \faIcon{sitemap}\hspace{0.2em}Network diagrams and system info
    \item \faIcon{question-circle}\hspace{0.2em}Discussion questions per phase
    \item \faIcon{book}\hspace{0.2em}Reference documents (IR plans)
    \item \faIcon{users}\hspace{0.2em}Participant roles and ground rules
\end{enumerate}

\section{\textcolor{accent}{\faIcon{users}}\hspace{0.4em}Key Participants}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{user-tag}\hspace{0.2em}\bfseries Role} & \textcolor{white}{\faIcon{tasks}\hspace{0.2em}\bfseries Responsibility} \\
\midrule
Facilitator & Guide discussion, present injects \\
OT Operations & Control room, field operations \\
OT Engineering & Technical system knowledge \\
IT Security & Detection/response expertise \\
Management & Escalation, business decisions \\
Safety & Safety implications review \\
\end{tabular}
\end{center}

\posterwarning{
\textcolor{warning}{\faIcon{exclamation-triangle}}\hspace{0.2em}\textbf{Include OT staff:} Tabletop exercises must include operations and engineering, not just IT security. Decisions about isolating control systems require input from those who understand operational consequences.
}

\section{\textcolor{accent}{\faIcon{crosshairs}}\hspace{0.4em}OT Scenario Categories}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{crosshairs}\hspace{0.2em}\bfseries Category} & \textcolor{white}{\faIcon{pencil-alt}\hspace{0.2em}\bfseries Example} \\
\midrule
Ransomware & IT to historian to HMI spread \\
Targeted APT & Eng. workstation, PLC logic mod \\
Insider threat & Disgruntled employee sabotage \\
Supply chain & Compromised vendor update \\
Safety system & Attacker disables SIS \\
Data integrity & Process values manipulated \\
\end{tabular}
\end{center}

\subsection{\textcolor{accent}{\faIcon{list-ol}}\hspace{0.3em}Scenario Progression}

\begin{enumerate}
    \item \faIcon{bell}\hspace{0.2em}\textbf{Initial detection} -- SOC alerts on suspicious activity
    \item \faIcon{search}\hspace{0.2em}\textbf{Investigation} -- Determine scope, affected systems
    \item \faIcon{level-up-alt}\hspace{0.2em}\textbf{Escalation} -- IT/OT coordination, management notify
    \item \faIcon{hand-paper}\hspace{0.2em}\textbf{Containment} -- Isolation decisions, operational impact
    \item \faIcon{undo}\hspace{0.2em}\textbf{Recovery} -- Eradicate threat, restore systems, verify integrity
    \item \faIcon{lightbulb}\hspace{0.2em}\textbf{Post-incident} -- Lessons learned, improvements
\end{enumerate}

\section{\textcolor{accent}{\faIcon{chalkboard-teacher}}\hspace{0.4em}Facilitation Tips}

\begin{itemize}
    \item \textcolor{info}{\faIcon{balance-scale}}\hspace{0.2em}\textbf{Stay neutral} -- Don't lead to ``correct'' answers
    \item \textcolor{success}{\faIcon{comments}}\hspace{0.2em}\textbf{Encourage participation} -- Draw out quieter members
    \item \textcolor{accent}{\faIcon{pencil-alt}}\hspace{0.2em}\textbf{Capture insights} -- Assign note-taker for gaps
    \item \textcolor{danger}{\faIcon{ban}}\hspace{0.2em}\textbf{No blame} -- Focus on process improvement
\end{itemize}

\section{\textcolor{accent}{\faIcon{file-alt}}\hspace{0.4em}After-Action Review}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{file-alt}\hspace{0.2em}\bfseries Section} & \textcolor{white}{\faIcon{align-left}\hspace{0.2em}\bfseries Content} \\
\midrule
Summary & Key findings, recommendations \\
Overview & Objectives, scenario, participants \\
Strengths & Validated capabilities \\
Improvements & Gaps with specific examples \\
Actions & Prioritized items with owners \\
\end{tabular}
\end{center}

\subsection{\textcolor{accent}{\faIcon{calendar-alt}}\hspace{0.3em}Exercise Frequency}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{1.5cm}p{2.5cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{list-ul}\hspace{0.2em}\bfseries Type} & \textcolor{white}{\faIcon{calendar-alt}\hspace{0.2em}\bfseries Freq.} & \textcolor{white}{\faIcon{bullseye}\hspace{0.2em}\bfseries Purpose} \\
\midrule
Basic tabletop & Quarterly & Awareness, test updates \\
Cross-functional & Semi-annual & IT/OT coordination \\
Executive & Annually & Strategic decisions \\
Full-scale & 2--3 years & End-to-end validation \\
\end{tabular}
\end{center}

\postertip{
Tabletop exercises are the \textbf{safest way to test OT incident response}. Include both IT security and OT operations/engineering staff. Use realistic scenarios based on actual industry threats. Progress through all IR phases: detection through recovery. The real value is in the \textbf{follow-through}---track action items to completion and update plans based on findings. CISA provides ready-made exercise packages for critical infrastructure.
}

\end{multicols}

\end{document}
