% ============================================================================
%  OT Backup and Recovery - Poster / Cheat Sheet
% ============================================================================

\documentclass[9pt,a4paper]{extarticle}
\usepackage{otsec-poster}
\usepackage{float}

\begin{document}

\makepostertitle
    {OT Backup and Recovery}
    {Strategies for Protecting and Restoring Industrial Control Systems}
    {Poster 850}
    {Matthias Niedermaier}

\begin{multicols}{2}

\section{\textcolor{accent}{\faIcon{info-circle}}\hspace{0.4em}Overview}

Unlike IT where data is the primary asset, OT must protect and restore \textbf{control logic, configurations, and operational state}---often for systems that cannot tolerate extended downtime. Without proper backups, \textcolor{danger}{\faIcon{exclamation-triangle}}\hspace{0.2em}ransomware or failures can mean weeks recreating control logic from scratch.

\section{\textcolor{accent}{\faIcon{columns}}\hspace{0.4em}OT vs IT Backup}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2cm}p{2cm}p{2.2cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{exchange-alt}\hspace{0.2em}\bfseries Aspect} & \textcolor{white}{\faIcon{desktop}\hspace{0.2em}\bfseries IT} & \textcolor{white}{\faIcon{cogs}\hspace{0.2em}\bfseries OT} \\
\midrule
Primary asset & Data files & Control logic, configs \\
Frequency & Daily/hourly & After changes \\
Restore window & Hours OK & Minutes critical \\
Agents & Standard & Often not possible \\
Network access & Always connected & Air-gapped common \\
Validation & File integrity & Functional testing \\
\end{tabular}
\end{center}

\section{\textcolor{accent}{\faIcon{hdd}}\hspace{0.4em}What to Back Up}

\posterinfo{
\textbf{Critical OT assets:} PLC/RTU programs (ladder logic, function blocks), HMI projects (graphics, tags, alarms), SCADA configurations, DCS controller configs, safety system logic, network device configs, and engineering workstation project files.
}

\subsection{\textcolor{accent}{\faIcon{database}}\hspace{0.3em}Operational Data}

\begin{itemize}
    \item \faIcon{chart-line}\hspace{0.2em}Historian databases (process data, trends)
    \item \faIcon{bell}\hspace{0.2em}Alarm and event logs
    \item \faIcon{industry}\hspace{0.2em}Batch records and production data
    \item \faIcon{clipboard-list}\hspace{0.2em}Audit trails and access logs
\end{itemize}

\subsection{\textcolor{accent}{\faIcon{file-alt}}\hspace{0.3em}Supporting Documentation}

\begin{itemize}
    \item \faIcon{project-diagram}\hspace{0.2em}Network diagrams and IP assignments
    \item \faIcon{plug}\hspace{0.2em}I/O lists and wiring documentation
    \item \faIcon{sliders-h}\hspace{0.2em}Setpoint and tuning parameters
    \item \faIcon{file-alt}\hspace{0.2em}Recovery procedures
\end{itemize}

\section{\textcolor{accent}{\faIcon{clock}}\hspace{0.4em}Backup Strategies}

\postersuccess{
\textbf{Change-triggered backups:} Perform a full backup immediately before and after any change to control logic. This creates a \textcolor{success}{\faIcon{check-circle}}\hspace{0.2em}known-good state and documents what changed.
}

\subsection{\textcolor{accent}{\faIcon{calendar-alt}}\hspace{0.3em}Scheduled Backups}

Even without changes, periodic backups catch:
\begin{itemize}
    \item \faIcon{user-secret}\hspace{0.2em}Unauthorized modifications
    \item \faIcon{sliders-h}\hspace{0.2em}Runtime parameter drift
    \item \faIcon{exclamation-triangle}\hspace{0.2em}Configuration corruption
    \item \faIcon{balance-scale}\hspace{0.2em}Comparison baselines for integrity monitoring
\end{itemize}

\subsection{\textcolor{accent}{\faIcon{download}}\hspace{0.3em}Backup Methods}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{1.8cm}p{2.2cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{tools}\hspace{0.2em}\bfseries Method} & \textcolor{white}{\faIcon{thumbs-up}\hspace{0.2em}\bfseries Advantage} & \textcolor{white}{\faIcon{thumbs-down}\hspace{0.2em}\bfseries Disadvantage} \\
\midrule
Online (running) & No downtime & May miss state \\
Offline (stopped) & Complete capture & Requires outage \\
Memory card & Direct, complete & Physical access \\
Network upload & Remote capable & Protocol limits \\
Disk image & Bare-metal restore & Large size \\
\end{tabular}
\end{center}

\section{\textcolor{accent}{\faIcon{lock}}\hspace{0.4em}Storage and Protection}

\posterdanger{
\textbf{Backups must be stored separately from production.} If ransomware encrypts your OT network, it should not reach your backups.
}

\subsection{\textcolor{accent}{\faIcon{layer-group}}\hspace{0.3em}3-2-1 Rule for OT}

\begin{itemize}
    \item \faIcon{copy}\hspace{0.2em}\textbf{3} copies of critical data
    \item \faIcon{hdd}\hspace{0.2em}\textbf{2} different storage media/types
    \item \faIcon{map-marker-alt}\hspace{0.2em}\textbf{1} copy offsite or air-gapped
    \item \faIcon{lock}\hspace{0.2em}Plus: at least one copy \textbf{offline} and immune to network attacks
\end{itemize}

\subsection{\textcolor{accent}{\faIcon{shield-alt}}\hspace{0.3em}Backup Security}

\begin{itemize}
    \item \faIcon{lock}\hspace{0.2em}\textbf{Encryption} -- AES-256 at rest and in transit
    \item \faIcon{user-lock}\hspace{0.2em}\textbf{Access control} -- Limit who can read/write/delete
    \item \faIcon{fingerprint}\hspace{0.2em}\textbf{Integrity verification} -- Hash validation, checksums
    \item \faIcon{database}\hspace{0.2em}\textbf{Immutable storage} -- Write-once or immutable cloud
    \item \faIcon{history}\hspace{0.2em}\textbf{Version retention} -- Keep multiple generations
\end{itemize}

\posterwarning{
\textbf{\textcolor{warning}{\faIcon{key}}\hspace{0.2em}Key management:} Encrypted backups are useless without decryption keys. Never store keys on the same media. Maintain offline key copies in physically secure storage. Include decryption in recovery drills.
}

\section{\textcolor{accent}{\faIcon{undo}}\hspace{0.4em}Recovery Planning}

\subsection{\textcolor{accent}{\faIcon{clock}}\hspace{0.3em}Recovery Time Objectives}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{1.5cm}p{2.5cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{server}\hspace{0.2em}\bfseries System} & \textcolor{white}{\faIcon{stopwatch}\hspace{0.2em}\bfseries RTO} & \textcolor{white}{\faIcon{flag}\hspace{0.2em}\bfseries Priority} \\
\midrule
Safety systems & Minutes & Critical \\
Critical PLCs & 1--4 hours & High \\
HMI/SCADA & 4--8 hours & High \\
Historians & 24 hours & Medium \\
Eng. stations & Days & Low \\
\end{tabular}
\end{center}

\section{\textcolor{accent}{\faIcon{vial}}\hspace{0.4em}Testing and Validation}

\posterdanger{
\textbf{Untested backups are not backups.} Many organizations discover their backups are corrupt, incomplete, or unusable only when needed most.
}

\subsection{\textcolor{accent}{\faIcon{calendar-check}}\hspace{0.3em}Testing Schedule}

\begin{itemize}
    \item \faIcon{robot}\hspace{0.2em}\textbf{Every backup} -- Automated verification (hash, completeness)
    \item \faIcon{calendar-alt}\hspace{0.2em}\textbf{Quarterly} -- Procedure review and tabletop exercise
    \item \faIcon{flask}\hspace{0.2em}\textbf{Semi-annually} -- Lab restoration test
    \item \faIcon{hard-hat}\hspace{0.2em}\textbf{Annually} -- Full recovery drill during maintenance
\end{itemize}

\postertip{
Back up PLCs, HMIs, network devices, and documentation. Use change-triggered backups. Follow the 3-2-1 rule with at least one offline copy. Test recovery regularly.
}

\end{multicols}

\end{document}
