% ============================================================================
%  OT Incident Response - Poster / Cheat Sheet
% ============================================================================

\documentclass[9pt,a4paper]{extarticle}
\usepackage{otsec-poster}
\usepackage{float}

\begin{document}

\makepostertitle
    {OT Incident Response}
    {Responding to Security Incidents in Industrial Environments}
    {Poster 600}
    {Matthias Niedermaier}

\begin{multicols}{2}

\section{\textcolor{accent}{\faIcon{info-circle}}\hspace{0.4em}Overview}

OT incident response requires different approaches than IT. The presence of \textbf{physical processes, safety systems, and operational constraints} means standard IR playbooks may be inappropriate or dangerous.

\posterdanger{
\textbf{\textcolor{danger}{\faIcon{fire-extinguisher}}\hspace{0.2em}Standard IT IR actions can cause harm in OT:} Disconnecting a PLC or rebooting a control system can stop production or create safety hazards. Isolate, reimage, restore---these common IT actions require extreme caution in OT.
}

\begin{center}
\begin{tikzpicture}[
    phase/.style={rectangle, draw=#1!50, thick, fill=#1!10, rounded corners=2pt, minimum height=0.4cm, minimum width=1.8cm, align=center, font=\scriptsize\bfseries},
    phase/.default={otaccent},
    arr/.style={->, thick, otaccent!60},
]
    \node[phase=otinfo] (p1) at (0,0) {Prepare};
    \node[phase=otwarning, right=4pt of p1] (p2) {Detect};
    \node[phase=otdanger, right=4pt of p2] (p3) {Contain};
    \node[phase=otsuccess, below=5pt of p3] (p4) {Eradicate};
    \node[phase=otaccent, below=5pt of p2] (p5) {Recover};
    \node[phase=otprimary, below=5pt of p1] (p6) {Lessons};

    \draw[arr] (p1) -- (p2);
    \draw[arr] (p2) -- (p3);
    \draw[arr] (p3) -- (p4);
    \draw[arr] (p4) -- (p5);
    \draw[arr] (p5) -- (p6);
    \draw[arr, dashed] (p6) -- (p1);
\end{tikzpicture}
\end{center}

\section{\textcolor{accent}{\faIcon{columns}}\hspace{0.4em}IT vs OT Response}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{1.8cm}p{2.2cm}p{2.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{columns}\hspace{0.2em}\bfseries Aspect} & \textcolor{white}{\faIcon{desktop}\hspace{0.2em}\bfseries IT} & \textcolor{white}{\faIcon{industry}\hspace{0.2em}\bfseries OT} \\
\midrule
Goal & Protect data & Safe operations \\
Isolation & Disconnect now & May not be possible \\
Restart & Common fix & Can disrupt process \\
Forensics & Image drives & Often no storage \\
Timing & Business hours & 24/7 operations \\
Expertise & IT security & Process knowledge \\
\end{tabular}
\end{center}

\section{\textcolor{accent}{\faIcon{list-ol}}\hspace{0.4em}IR Phases}

\subsection{\textcolor{accent}{\faIcon{clipboard-check}}\hspace{0.3em}1. Preparation}

\begin{itemize}
    \item \faIcon{users}\hspace{0.2em}Cross-trained team (IT security + OT engineering)
    \item \faIcon{book}\hspace{0.2em}OT-specific playbooks reviewed by operations
    \item \faIcon{sitemap}\hspace{0.2em}Network diagrams and asset inventory available
    \item \faIcon{hdd}\hspace{0.2em}Backup configurations for critical devices
    \item \faIcon{handshake}\hspace{0.2em}Vendor relationships established
    \item \textcolor{danger}{\faIcon{hard-hat}}\hspace{0.2em}Safe shutdown procedures documented
\end{itemize}

\subsection{\textcolor{accent}{\faIcon{bell}}\hspace{0.3em}2. Detection and Analysis}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{search}\hspace{0.2em}\bfseries Source} & \textcolor{white}{\faIcon{bell}\hspace{0.2em}\bfseries Detection Method} \\
\midrule
OT monitoring & Anomalous traffic, unauthorized cmds \\
Process anomalies & Unexpected behavior from operators \\
IT alerts & Threats moving toward OT network \\
Vendor notify & Vulnerability/compromise advisories \\
\end{tabular}
\end{center}

\subsection{\textcolor{accent}{\faIcon{hand-paper}}\hspace{0.3em}3. Containment}

\posterwarning{
\textbf{\textcolor{warning}{\faIcon{fire-extinguisher}}\hspace{0.2em}Containment requires extreme caution:} Consult operations before any network changes. Understand process dependencies before isolation. Consider safety implications. Document everything---changes may need reversal.
}

\textcolor{warning}{\faIcon{fire-extinguisher}}\hspace{0.2em}Containment options (least to most disruptive):

\begin{enumerate}
    \item \faIcon{eye}\hspace{0.2em}\textbf{Monitor only} -- Observe while gathering intel
    \item \faIcon{shield-alt}\hspace{0.2em}\textbf{Block at firewall} -- Stop specific traffic
    \item \faIcon{project-diagram}\hspace{0.2em}\textbf{Segment} -- Isolate zone, maintain internal ops
    \item \faIcon{power-off}\hspace{0.2em}\textbf{Controlled shutdown} -- Safe process stop
    \item \textcolor{danger}{\faIcon{hard-hat}}\hspace{0.2em}\textbf{Emergency stop} -- Only for imminent safety threat
\end{enumerate}

\subsection{\textcolor{accent}{\faIcon{undo}}\hspace{0.3em}4. Eradication and Recovery}

\begin{itemize}
    \item \textcolor{warning}{\faIcon{fire-extinguisher}}\hspace{0.2em}\textbf{Known-good restore} -- Use verified config backups
    \item \faIcon{phone}\hspace{0.2em}\textbf{Vendor involvement} -- May need vendor support
    \item \faIcon{layer-group}\hspace{0.2em}\textbf{Staged recovery} -- Bring systems back incrementally
    \item \faIcon{check-double}\hspace{0.2em}\textbf{Verification} -- Confirm correct process operation
    \item \faIcon{broadcast-tower}\hspace{0.2em}\textbf{Enhanced monitoring} -- Watch for reinfection
\end{itemize}

\section{\textcolor{accent}{\faIcon{hard-hat}}\hspace{0.4em}Safety First}

\posterdanger{
\textbf{\textcolor{danger}{\faIcon{hard-hat}}\hspace{0.2em}Safety always takes precedence over security:}
\begin{itemize}
    \item \textcolor{danger}{\faIcon{hard-hat}}\hspace{0.2em}Never disable safety instrumented systems
    \item \textcolor{danger}{\faIcon{hard-hat}}\hspace{0.2em}Ensure safe state before any remediation
    \item \textcolor{danger}{\faIcon{hard-hat}}\hspace{0.2em}Involve process safety personnel in decisions
    \item \textcolor{danger}{\faIcon{hard-hat}}\hspace{0.2em}Document safety implications of all actions
\end{itemize}
}

\subsection{\textcolor{accent}{\faIcon{exclamation-triangle}}\hspace{0.3em}Operational Constraints}

Consider before taking action:

\begin{itemize}
    \item \faIcon{cog}\hspace{0.2em}\textbf{Process state} -- Can the process be safely interrupted?
    \item \faIcon{industry}\hspace{0.2em}\textbf{Production} -- Critical batches or orders in progress?
    \item \faIcon{calendar-alt}\hspace{0.2em}\textbf{Maintenance windows} -- When can changes be made?
    \item \faIcon{users}\hspace{0.2em}\textbf{Staffing} -- Are qualified operators available?
    \item \faIcon{link}\hspace{0.2em}\textbf{Dependencies} -- What else will be affected?
\end{itemize}

\section{\textcolor{accent}{\faIcon{users}}\hspace{0.4em}Team Structure}

\begin{center}
\rowcolors{2}{lightgray}{white}
\begin{tabular}{p{2.2cm}p{4.3cm}}
\rowcolor{primary}
\textcolor{white}{\faIcon{user-tag}\hspace{0.2em}\bfseries Role} & \textcolor{white}{\faIcon{tasks}\hspace{0.2em}\bfseries Responsibility} \\
\midrule
IT Security & Malware analysis, network forensics \\
OT Engineering & Process knowledge, control systems \\
Operations & Current process state, safe shutdown \\
Safety & Risk assessment, safety expertise \\
Management & Decision authority, communications \\
Vendors & System-specific expertise \\
\end{tabular}
\end{center}

\section{\textcolor{accent}{\faIcon{lightbulb}}\hspace{0.4em}Post-Incident}

\postersuccess{
\textbf{Lessons learned must cover:} What happened (root cause, timeline). What worked (effective detection/response). What failed (monitoring gaps, procedure issues). Improvements (specific actions to prevent recurrence). Update IR procedures based on findings.
}

\postertip{
Prepare OT-specific playbooks \textbf{before} an incident. Cross-train IT security and OT engineering teams. Safety always overrides security---never take actions that could create unsafe conditions. Use graduated containment: monitor first, isolate only when necessary and safe. Establish out-of-band communications and vendor contacts in advance. Every minute spent on preparation saves hours during an actual incident.
}

\end{multicols}

\end{document}
